CL-PFU: A suite of R packages for energy conversion chain analysis

### 2023-08-17 paper.Rmd
Working on PFUAggDatabase pipeline
Finished both pipelines.

<!-- Define some macros to use later -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaweeb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

We created a suite of `R` packages 
to assist SEA practitioners to analyze 
energy movement through society. 
The new packages enable analysis of any country in the world
across timespans of decades or longer.
In short, the packages enable, for the first time, scalable SEA.
We used the new packages to create the CL-PFU database,^["CL-PFU" 
is an initialism for "Country-Level
Primary, Final, and Useful."
The CL-PFU database is "multi-regional" in the sense that
it contains many countries as well as 
continental and world aggregations.
Technically speaking, 
we create [matsindf][link-matsindf] data frames, 
not SQL or similar databases.] 
a new resource for the SEA community [@Marshall:2023aa].

@@

```{r}
# (1) Purchase, download, and store IEA data in correct location.
# (2) Download FAO and ILO data for muscle work calculations.
source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
# (3) Run the PFUDatabase pipeline
targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
                  reporter = "summary")

# (4) Run the PFUAggDatabase pipeline
# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
```

Create a CL-PFU database from the GHA and ZAF data contained in the IEATools package.   
Create the folder structure for the CL-PFU database.   
Store all necessary data in this repository.   
Execute the commands to produce the database, 
albeit for only two countries and two years.

>>>>

<!-- Define some macros to use later -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaweeb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

We created a suite of `R` packages 
to assist SEA practitioners to analyze 
energy movement through society. 
The new packages enable analysis of any country in the world
across timespans of decades or longer.
In short, the packages enable, for the first time, scalable SEA.
We used the new packages to create the CL-PFU database,^["CL-PFU" 
is an initialism for "Country-Level
Primary, Final, and Useful."
The CL-PFU database contains many countries as well as 
continental and world aggregations.
Technically speaking, 
we create [matsindf][link-matsindf] data frames
stored as `[pins][link-pins]` on a pinboard, 
not SQL or similar databases.] 
a new resource for the SEA community [@Marshall:2023aa].

@@

```{r, eval=FALSE}
# (1) Purchase, download, and store IEA data in correct location.
# (2) Download FAO and ILO data for muscle work calculations
# via the included script.
# This download is needed only once and
# the results are now stored in the repository. 
source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
```

```{r, eval=FALSE}
# (3) Run the PFUDatabase pipeline
targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
# (4) Run the PFUAggDatabase pipeline
# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"), callr_function = NULL)
targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
                         workers = 4)
```

```{r}
# (5) Review results using the saved pins
# Establish the pinboard
pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
  pins::board_folder(versioned = TRUE)
# Look at the psut data frame
# created by the PFUDatabase pipeline
psut <- pinboard |>
  pins::pin_read("psut", version = "20230811T122618Z-3b7b3")
head(psut)
# Each column of the psut data frame contains matrices.
# The data are stored as sparse matrices,
# but easier to see as regular matrices.
psut$R[[1]] |> 
  as.matrix()

# Look at the aggregate PFU efficiency data frame
# created by the PFUAggDatabase pipeline
agg_eta_pfu <- pinboard |> 
  pins::pin_read("agg_eta_pfu", version = "")
head(agg_eta_pfu)
```
